---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- [![License](https://img.shields.io/badge/License-Apache%202.0-blue.svg)](https://opensource.org/licenses/Apache-2.0) -->
<!-- [![R build status](https://github.com/bcgov/dipr/workflows/R-CMD-check/badge.svg)](https://github.com/bcgov/dipr) -->
<!-- <a id="devex-badge" rel="Exploration" href="https://github.com/BCDevExchange/assets/blob/master/README.md"><img alt="Being designed and built, but in the lab. May change, disappear, or be buggy." style="border-width:0" src="https://assets.bcdevexchange.org/images/badges/exploration.svg" title="Being designed and built, but in the lab. May change, disappear, or be buggy." /></a> -->

<!-- badges: end -->

# dipr

The `dipr` package is an R package that loads and provides means of caching of `.dat.gz` files. Functions are also provided to convert from `.dat.gz` files to Apache Arrow formats.

## Installation
While this package is intended for us within the SRE you can install the package from GitHub:

```{r, eval=FALSE}
if(!requireNamespace("remotes")) install.packages("remotes")
remotes::install_github("bcgov/dipr")
```



## Installation inside the SRE
Installation of `dipr` is slightly different than typical R package installation. To begin with you will need to clone the dipr repo. To do this you need access to the SRE instance of GitLab. 

Go to this website: https://projectsc.popdata.bc.ca/

Scroll down and click on "Sign in with Pop Data Sso"

You will need to do this step everytime you log into the SRE.

Next go to this site: https://projectsc.popdata.bc.ca/profile/personal_access_tokens

Create a new Personal Access Token enabling each scope and typing in your username. This should create a new personal access token. Those are your "git credentials". Save those on your "U" drive along with your user name. 

### Clone from RStudio

Open RStudio and go to 
File -> New Project -> Version Control -> Git

In the field `Repository URL` section type:
https://projectsc.popdata.bc.ca/shares/dipr

In the field `Create project as subdirectory of` browse to where you want to clone this repo. I'd recommend putting this in a personal but open to other team members folder.


### Clone from bash

Navigate to where you would like to clone the repo.

```
git clone https://projectsc.popdata.bc.ca/shares/dipr
```

### Installation using devtools
This will clone the repo and you can install from there. Inside the repo will be an `dipr.Rproj`. Double clicking on that will open RStudio. Next run the following code:

```r
devtools::install()

```

Once you have the package installed you can use `dipr::dipr_update` to update. 


## Usage
The `dipr` package comes with some sample _fake_ data to illustrate the usage of the package and facilitate unit tests. Specifically, the package provides a fake data dictionary and a compressed data file. Example fake data can be seen like this:
```{r}
library(dipr)
dipr_examples()
```

Individual paths can be extracted like this:
```{r}
data_dict_path <- dipr_example("starwars-dict.txt")
dict <- read.table(data_dict_path)
dict
dat_path <- dipr_example("starwars-fwf.dat.gz")
```

The main function in the package is `read_dat()`:
```{r}
raw <- read_dat(data_path = dat_path,
                data_dict = dict,
                use_cache = FALSE)
head(raw)
```

### Data Dictionary
Data Dictionaries need to be in a data frame format that includes 'name`, 'start', 'stop', and 'length' columns (in any order):

```r
> dict
# A tibble: 64 x 4
   start  stop length name                      
   <dbl> <dbl>  <dbl> <chr>   
```


### Caching
The approach to caching is for `read_dat` to create a `.dipr` folder in your project directory and store all cached files there. This works well if you are working primarily alone. If, however, you are working collaboratively, you can also set the cache data directory using the `DIPR_CACHE_PATH` environment variable. You can set this environment variable in your .Renviron file. Other users can do the same and then you are will be drawing cached data from the same source. 

### cache_type
Users can specify the format of the cached files via the `cache_type` argument in `read_dat`. The default option is currently the `fst` format from the `fst` package because it is the fastest option in R. Support for `parquet` is still very developmental and is not recommended for widespread usage. 


## Project Status
Under development

## License


    Copyright 2019 Province of British Columbia

    Licensed under the Apache License, Version 2.0 (the "License");
    you may not use this file except in compliance with the License.
    You may obtain a copy of the License at 

       http://www.apache.org/licenses/LICENSE-2.0

    Unless required by applicable law or agreed to in writing, software
    distributed under the License is distributed on an "AS IS" BASIS,
    WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
    See the License for the specific language governing permissions and
    limitations under the License.
   
